EinsteinPy: Python for General Relativity\footnote{Released on July, 15th, 2019}

### 2020-01-18 authors.tex
Seaparate Authors.tx

%%
%% Beginning of file 'paper.tex'
%%
%% Modified 2018 January
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.2 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
%% All of these external packages should already be present in the modern TeX 
%% distributions.  If not they can also be obtained at www.ctan.org.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.2
\documentclass{aastex62}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be envoked like this:
%%
%% \documentclass[argument]{aastex62}
%% 
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Preferred style for Research Notes which are by design 
%%                lacking an abstract and brief. DO NOT use \begin{abstract}
%%                and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can envoke to allow other stylistic
%% actions. The available options are:
%%
%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%  tighten      : Makes baselineskip slightly smaller, only works with 
%%                 the twocolumn substyle.
%%  times        : uses times font instead of the default
%%  linenumbers  : turn on lineno package.
%%  trackchanges : required to see the revision mark up and print its output
%%  longauthor   : Do not use the more compressed footnote style (default) for 
%%                 the author/collaboration/affiliations. Instead print all
%%                 affiliation information after each name. Creates a much
%%                 long author list but may be desirable for short author papers
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex62}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

%% Tells LaTeX to search for image files in the 
%% current directory as well as in the figures/ folder.
\graphicspath{{./}{figures/}}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{January 1, 2018}
\revised{January 7, 2018}
\accepted{\today}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{ApJ}

%% Mark up commands to limit the number of authors on the front page.
%% Note that in AASTeX v6.2 a \collaboration call (see below) counts as
%% an author in this case.
%
%\AuthorCollaborationLimit=3
%
%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
%% collaboration" on the front page of this example manuscript.
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.
%\usepackage{minted}
\usepackage{listings}

\begin{document}

\title{EinsteinPy: Python for General Relativity\footnote{Released on July, 15th, 2019}}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.2 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affilations. The change was necessary to implement 
%% autoindexing of affilations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% arguement which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.2 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.2 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

\author[0000-0002-0870-4665]{Shreyas Bapat}
\affiliation{School of Computing and Electrical Engineering, Indian Institute of Technology Mandi, India}

\author{Bhavya Bhatt}
\affiliation{School of Computing and Electrical Engineering, Indian Institute of Technology Mandi, India}

\author{Ritwik Saha}
\affiliation{School of Computing and Electrical Engineering, Indian Institute of Technology Mandi, India}

\author[0000-0003-1021-3336]{Priyanshu Khandelwal}
\affiliation{School of Computing and Electrical Engineering, Indian Institute of Technology Mandi, India}
\collaboration{(EinsteinPy Core Development Committee)}

\author[0000-0002-2268-9772]{Sofía Ortín Vela}
\affil{Department of Theoretical Physics, University of Zaragoza, Spain}

@@

\author[0000-0001-7234-6148]{Varun Singh}
\affiliation{School of Computing and Electrical Engineering, Indian Institute of Technology Mandi, India}

\author[0000-0002-1900-3338]{Shilpi Jain}
\affiliation{Department of Earth Sciences, Indian Institute of Technology Roorkee, India}

@@

\author[0000-0002-5198-9010]{Manvi Gupta}
\affiliation{School of Computing and Electrical Engineering, Indian Institute of Technology Mandi, India}

\author{Tushar Tyagi}
\affiliation{School of Computing and Electrical Engineering, Indian Institute of Technology Mandi, India}

\author[0000-0003-1838-8368]{Tanmay Rustagi}
\affiliation{School of Computing and Electrical Engineering, Indian Institute of Technology Mandi, India}

\author[0000-0002-0757-2883]{Abhijeet Manhas}
\affiliation{School of Computing and Electrical Engineering, Indian Institute of Technology Mandi, India}

@@

\author[0000-0001-8302-1584]{Sashank Mishra}
\affiliation{Department of Information Technology, Indian Institute of Information Technology Allahabad}
\collaboration{(EinsteinPy Package Contributors)}
%% Note that the \and command from previous versions of AASTeX is now
%% depreciated in this version as it is no longer necessary. AASTeX 
%% automatically takes care of all commas and "and"s between authors names.

%% AASTeX 6.2 has the new \collaboration and \nocollaboration commands to
%% provide the collaboration status of a group of authors. These commands 
%% can be used either before or after the list of corresponding authors. The
%% argument for \collaboration is the collaboration identifier. Authors are
%% encouraged to surround collaboration identifiers with ()s. The 
%% \nocollaboration command takes no argument and exists to indicate that
%% the nearby authors are not part of surrounding collaborations.

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}

This paper presents EinsteinPy (version 0.2), a community-developed Python
package for gravitational and relativistic astrophysics. Python is a free, easy to use high level programming language which has seen a huge expansion in the number of its users and developers in recent years. Specifically, a lot of recent studies show that the use of Python in Astrophysics and in general physics has increased exponentially. Many great frameworks came as Python packages which provide a very high level of abstraction over the dirty nitty-gritty of complex algorithms and provide an easy to use interfaceand pleasing user experience. One such example is Keras - framework for deep learning which has made deep learning so easy that a person with zero programming knowledge can also train a neural network classifier. This example really demonstrates the power of abstraction which is achievable in Python. The aim of the EinsteinPy is no different and is developed keeping in mind the state of a theoretical gravitational physicist with a little or no background in computer programming and trying to work in the field of numerical relativity or trying to use simulations in their research. Currently EinsteinPy supports simulation of time-like and null geodesics and calculate trajectories in different background geometries some of which are Schwarzschild, Kerr and KerrNewmann along with coordinate inter-conversion pipeline. It has a partially developed pipeline for plotting and visualization with dependencies on libraries like plotly, matplotlib etc. One of the unique feature of EinsteinPy is a sufficiently developed symbolic tensor manipulation utilites which is a great tool in itself for teaching yourself tensor algebra which for many beginner students can be overwhelmingly tricky. Currently EinsteinPy also provide few utility functions for hypersurface embedding of Schwarzschild spacetime which further will be extended to model gravitational lensing simulation. The current version of the library is in a state that can be used by any serious student of general relativity trying to get essence of this beautiful subject but is somewhere lost in the heavy mathematical formalism of the subject. EinsteinPy provides such students to really see through the equations and visualize whats really happening.

\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{gravitational physics, astrophysics --- 
simulations --- black holes --- gravitational waves}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.
%%
%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

\section{Introduction} \label{sec:intro}

The relativistic theory of gravity paved its way to the ground in 1915, when Einstein published his paper on the general theory of relativity. This elegant and rigorous framework  was a generalized version of gravity-free theory - special relativity, which he published earlier in 1905 and since then the whole physics community was against the bold ideas of the young genius and resisted the same in the beginning, as all people were worried that if the hypothesis stood true, complete notion of how they perceived space and time would get altered. Sooner,  they started to realize and conceive the true depth in the formalism of general relativity and its ability to explain the fundamental laws of nature. It took a huge span of time for the acceptability of general theory of relativity as the "theory of gravity" in the classical regime. All the attempts to formulate a quantum theory of gravity in some way or the other borrow its ideas from General Relativity.
Almost a hundred years after Einstein wrote down the equations of General Relativity, solutions of the Einstein equation remain extremely difficult to find beyond those which exhibit significant symmetries. The central problem then and now remains to be the solutions of Einstein's field equation. Many times we can study the behaviour of solutions under a high degree of symmetry considerations and could even solve analytically for highly unrealistic systems but solving it for problems relevant to astrophysical and gravitational physics research. It still remains a big daunting question on how to get around the problem of solving these field equations. This question is so profound that it has a separate profound field of research which goes with the name of numerical relativity which attempts to use computation science and programming to numerically obtain solutions of the equations (which would be the background geometry) for some turbulent region where most of the interesting dynamics are happening (as we can not solve for an infinitely large grid due to restrictions imposed on us by space and time complexity and computability).\\
Numerical methods have been used to solve the Einstein equation for many decades, but the past decade has seen tremendous advances and adding to that, the interest of the community grew when the field found applications in areas of radio astronomy, cosmology, signal processing, data mining. The field of numerical relativity is growing rampantly with the vast literature on algorithms, numerical methods and theoretical formulations (from basic 3+1 decomposition formulation to more sophisticated ones) with the development of robust and involved frameworks that provide a complete programming ecosystem and have proved to be essential tools for any numerical relativity researcher. Reflecting the other end of the research community stands the major section of theoretical physicists which counts the majority as a novice in programming. The head challenge is the fact that the usage of these frameworks demands heavy use of high-level programming languages like C, C++, Python etc. Though Python provides a vast room for abstractions no library existed that focused towards numerical relativity, eventually, the need of python library dedicated to general relativity became the propellant for the team and thus they made dauntless efforts to cover the space for this Python library. Since then, EinsteinPy has seen a lot of contributions from people all over the world and many "good to go" functionalities are already provided in this and previous versions. Like any other numerical relativity library, EinsteinPy is made to provide a set of tools which can make numerical computations for solving Einstein's field equations an easy job for anyone who does not want to deal with intricacies of the subject along with few, very basic but powerful functionalities that could be used by anyone who wants to learn the subject of general relativity 
with every minute detail .in general\footnote{More details on this are given in further coming sections}. In further coming section, we discuss with some of the features of the current as well previous version and the future plans which are yet to be implemented in the upcoming versions. Also, we describe few code snippets to explain the usage of the library on which more details can be found on the organisation website\footnote{https://einsteinpy.org/}
Software development and tools form a significant part of research in astronomy and moreover, the field itself has long supported the development of software tools for astronomical tasks.. These packages therefore typically provide some level of documentation and user support or training.
.
The development of the EinsteinPy core package began as a community-driven effort to standardize core functionality for software in Python which renders a  user-friendly interface for supporting numerical relativity and relativistic astrophysics research. The library is as an attempt to provide programming and numerical environment for numerous numerical relativity problems like geodesics plotter, gravitational lensing and ray tracing, solving and simulating relativistic hydrodynamical equations, plotting of black hole event horizons, solving Einstein’s field equations and simulating various dynamical systems like binary merger etc.

At the same time, the scientific community contributing to the package has grown tremendously and an ecosystem of packages supporting or affiliated with the EinsteinPycore has developed. 
EinsteinPy provides an open-source and open-development core package and an ecosystem of affiliated 
packages that support numeric relativity functionality in the Python programming language. "Open development" describes a process where anybody with an internet connection can suggest changes to the source code and contribute their opinion when new features, bug fixes or other code changes, governance, or any other aspect of the development process is discussed. EinsteinPy core package is now a feature-rich library of sufficiently general tools and classes that support the development of more specialized code

In this paper, we describe the current status of the EinsteinPy core package and discuss goals for future development.
We start by describing the way EinsteinPy functions followed by the main software efforts developed by the EinsteinPy itself with proper detailed documentation of the current version release.We end with a short vision for the future of EinsteinPy to its contribution in the field of computational methodology in gravitational and relativistic Astrophysics.

\section{Units Handling}
EinsteinPy has dependencies on Astropy for handling units and user have to use appropriate Astropy units\footnote{more on astropy units can be referred from their official website documentation} while declaring the input data to various methods which are further used in the computation. 
\section{Data Types} \label{sec:style}
Since its inception,  EinsteinPy has aimed to provide is achieved by some set of data structures that are specifically designed to ease with a variety of gravitational physics problems. The most important and central quantity in all of the general relativity is the metric tensor $g_{\mu\nu}$ which defines the background geometry of spacetime on which all the dynamics happen. Informally the problem that packages of numerical relativity tend to solve is that we have a system (or matter field) defined by energy-momentum tensor $T_{\mu\nu}$ and we want to solve for the $g_{\mu\nu}$ such that it satisfies the following Einstein's field equation
\[
R_{\mu\nu} - \frac{1}{2}Rg_{\mu\nu} + \Lambda g_{\mu\nu} = \frac{8\pi G}{c^4}T_{\mu\nu}
\]where $R_{\mu\nu}$ is contracted Riemannian tensor or Ricci tensor, $R$ is Ricci scalar, $\Lambda$ is cosmological constant. All the above quantities are derivatives of metric tensor\footnote{more on this can be referred from any standard general relativity text} so in simple terms, on the left-hand side of the equation are terms involving $g_{\mu\nu}$ and on the right-hand side are the matter field terms which only characterize the distribution of matter.\\
EinsteinPy provides a core module: \texttt{metric} which has core data types of various metric tensors some which include Schwarzschild, Kerr, Kerr-Newman spacetime geometries(which are the exact solution of field equation for very specific matter distributions). The immediate functionalities which come along with this module are the solutions of geodesic equations. The module has coupled integration with \texttt{coordinates} module which deals with different coordinate system transformations\footnote{more on this in the 'coordinates' section}. Another core data types are in the \texttt{plotting} module which handles all the plotting, rendering, animation and simulation related methods and provides user-friendly interfacing of plots. One interesting feature of EinsteinPy is the support for symbolic tensor algebra which is extremely important in the formalism of general relativity and one of the first challenges faced by many beginners in their journey of learning the subject. EinsteinPy provides an excellent way to learn the math behind tensor manipulation and index gymnastics with the help of symbolic module: \texttt{symbolic}, on which more will be discussed in its section ahead.\\

Features of EinsteinPy and section division.
\textbf{Paper organisation}

			Organisation and Infrastructure
EinsteinPy equations and notations:
The purpose of this subsection is to clearly state the equations of motion that we have implemented and our conventions for completeness

Evolution Equation

EinsteinPy core package version 0.2
Integrators

Metric

Utils

Plotting

Constant

Coordinates

Geodesic

Bodies

Hypersurface

symbolic
Support of EinsteinPy
The EinsteinPy Project, as of the version 0.2release, does not receive any direct financial support for the development of EinsteinPy. Development of the software, all supporting materials, and community support are provided by individuals who work on EinsteinPy in their own personal time, by individuals or groups contributing to EinsteinPy as part of a research project.  ( A list of organizations that have contributed to EinsteinPy in this manner can be found in the Acknowledgments)

Release Cycle and Long-term Support
The EinsteinPy package is updated with significant changes and bug fixes in every release. The major releases contain new features or any significant changes, whereas the bugfix releases only contain fixes to code or documentation but no new features.

Following sections will give a breif overview of every core module and some of their important functionalities.
\subsection{tensor}
\subsection{metric}
\texttt{metric} module give users the freedom to define metric object which can be further used in the computations related to spacetime geometry. Let's see an small example of solving geodesic equation in Kerr spacetime and obtaining trajectories using EinsteinPy
\subsection{coordinates}
%\begin{minted}[linenos]{python}
%# importing modules
from astropy import units as u
import numpy as np
from einsteinpy.metric import Kerr
from einsteinpy.coordinates import BoyerLindquistDifferential

%# initializing parameters and coordinate system
%M = 1.989e30 * u.kg
%a = 0.3 * u.m
%BL_obj = BoyerLindquistDifferential(50e5 * u.km, np.pi / 2 * u.rad, np.pi * u.rad,
     %                               0 * u.km / u.s, 0 * u.rad / u.s, 0 * u.rad / u.s,
          %                          a)
%end_lambda = ((1 * u.year).to(u.s)).value / 930

%# Choosing stepsize for ODE solver to be 0.02 minutes
%stepsize = ((0.02 * u.min).to(u.s)).value

%# defining Kerr spacetime metric object
%obj = Kerr.from_coords(BL_obj, M)

%# calculates the trajectory
%ans = obj.calculate_trajectory(
    %end_lambda=end_lambda, OdeMethodKwargs={"stepsize": stepsize}, return_cartesian=True
%)
%\end{minted}
\subsection{plotting}
\subsection{symbolic}
General Relativity required heavy mathematical formulation to derive and describe various tensors attributing to various features essential for cogent description of a arbitrary space-time. We would describe the various quantities currently supported by this module while maintaining brevity. Each quantity has its own class, even scalars, which are treated as 0\textsuperscript{th} order tensors.

\subsubsection{BaseRelativityTensor}
This is a base class from which other tensors are derived. The user can use this class to create his/her own tensor while using the various in-built functions provided by this class. The class also maintains the indicial configuration of the tensor(contravariant and co-variant indices), calculates the free variables used in describing the tensor other than those which depicts the axis of space-time itself (e.g. t,x,y,z). The class has a function to convert the symbolic tensorial quantities into functions where actual numerical values can be substitued.

This class along with its parent class contains the following functions/properties.
\begin{itemize}
\item \texttt{order} (Property) : Returns the order of the tensor. 
\item \texttt{config} (Property) : Returns the arrangement of contravariant(upper)/covariant(lower) indices of the tensor. For example, 'ulll'.
\item \texttt{parent\_metric} (Property) : Returns the metric tensor of the space-time associated with the tensor, if present. Else returns \texttt{None}.
\item \texttt{tensor} (Function) : Returns the tensorial quantity contained within the class instance.
\item \texttt{subs} (Function) : Returns a new tensor upon substitution of current variables with new variables/values.
\item \texttt{simplify} (Function) : Returns a tensor contain simplified expressions from the current tensor.
\item \texttt{symbols} (Function) : Returns the symbols denoting each axis of the space-time.
\item \texttt{tensor\_lambdify} (Function) : Returns lambdified function of symbolic tensors. This means that the returned functions can accept numerical values and return numerical quantities.
\end{itemize}

\subsubsection{MetricTensor}
Metric Tensor describes the differential length elements required to measure distance in a curved(also applies to flat) space-time. It is a second order tensor and is fundamental to describe any space-time geometry. The \texttt{MetricTensor} class, being inherited from \texttt{BaseRelativityTensor}, inherits all its functions and also have some functions and constraints of its own. For example, metric tensor is bound to be a second order tensor.

The defined class functions are :
\begin{itemize}
\item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
\item \texttt{inv} : returns inverse of the metric
\end{itemize}

\subsubsection{ChristoffelSymbols}
Although Christoffel Symbols are belong to class of pseudo-tensors, it is inherited from \texttt{BaseRelativityTensor} to maintain homogenity within the module. Bein inherited from \texttt{BaseRelativityTensor}, it already supports the functions of its parent class and imposes some restrictions of its own, such as order of the tensor should be 3.

Instance of the class can be made by directly passing a 3\textsuperscript{rd} order tensor or can be obtained using the following classmethods
\begin{itemize}
\item \texttt{from\_metric} : Calculates christoffel symbols from a given instance of \texttt{MetricTensor} class.
\end{itemize}

The defined class function(s) are : 
\begin{itemize}
\item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
\end{itemize}

\subsubsection{RiemannCurvatureTensor}
This quantity is the most extensive measure of curvature of a sapce-time. It is inherited from \texttt{BaseRelativityTensor} and supports all the functions of its parent class.

Instance of the class can be made by directly passing a 4\textsuperscript{th} order tensor or can be obtained using the following classmethods
\begin{itemize}
\item \texttt{from\_metric} : Calculates Riemann Tensor from a given instance of \texttt{MetricTensor} class.
\item \texttt{from\_christoffels} : Calculates Riemann Tensor from a given instance of \texttt{ChristoffelSymbols} class.
\end{itemize}

The defined class function(s) are : 
\begin{itemize}
\item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
\end{itemize}

\subsubsection{RicciTensor}
This tensor is used in solving Eintein's  equation and basically contraction of Riemann Curvature Tensor.  It is inherited from \texttt{BaseRelativityTensor} and supports all the functions of its parent class.

Instance of the class can be made by directly passing a 2\textsuperscript{nd} order tensor or can be obtained using the following classmethods
\begin{itemize}
\item \texttt{from\_metric} : Calculates Ricci Tensor from a given instance of \texttt{MetricTensor} class.
\item \texttt{from\_christoffels} : Calculates Ricci Tensor from a given instance of \texttt{ChristoffelSymbols} class.
\item \texttt{from\_riemann} : Calculates Ricci Tensor from a given instance of \texttt{RiemannCurvatureTensor} class.
\end{itemize}

The defined class function(s) are : 
\begin{itemize}
\item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
\end{itemize}

\subsubsection{RicciScalar}
It is a scalar quantity obtained by contracting a given Ricci Tensor. It is inherited from \texttt{BaseRelativityTensor} and supports all the functions of its parent class. Being a scalar quantity, it does not support \texttt{change\_config} unlike other tensors as discussed above.

Instance of the class can be made by directly passing a symbolic expression or can be obtained using the following classmethods
\begin{itemize}
\item \texttt{from\_metric} : Calculates Ricci Scalar from a given instance of \texttt{MetricTensor} class.
\item \texttt{from\_christoffels} : Calculates Ricci Scalar from a given instance of \texttt{ChristoffelSymbols} class.
\item \texttt{from\_riemann} : Calculates Ricci Scalar from a given instance of \texttt{RiemannCurvatureTensor} class.
\item \texttt{from\_riccitensor} : Calculates Ricci Scalar from a given instance of \texttt{RicciTensor} class.
\end{itemize}

The defined class properties are : 
\begin{itemize}
\item \texttt{expr} : returns the scalar symbolic expression contained in the class. 
\end{itemize}

\subsubsection{Other Defined Quantities}
As the library is in heavy development, new quantities are being added as and when required. Some other quantities currently supported are listed below. The quantities are all inherited from \texttt{BaseRelativityTensor}, and a similar API (same as that of discussed above) has been implemented to maintain compatiblity and homegenity. The class names denote usual quantities relevant to General Relativity.

\begin{itemize}
\item \texttt{EinsteinTensor}
\item \texttt{StressEnergyMomentumTensor}
\item \texttt{WeylTensor}
\item \texttt{SchoutenTensor}
\end{itemize}

Further, various constants like $c$(speed of light), $G$(Gravitational constant) are defined in their symbolic forms, for use within the symbolic module.

\subsection{hypersurface}
This module provides with basic computational functions which are essential for modelling the space-like hypersurface for any spacetime geometry. Currently module has implementation for Schwarzschild geometry with conversion functions from Schwarzschild coordinates to 3-D spherical coordinates and some plotting utilities. The whole module is expected to extend for providing a superclass for implementations related to gravitationl lensing simulations and numerical calculations of null geodesics on hypersurfaces and deviation angles for systems which are relevant to relativistic astrophysical problems. A small code snippet for plotting the hypersurface is as follows:
%\begin{minted}[linenos]{python}
%# importing the modules
from einsteinpy.hypersurface import SchwarzschildEmbedding
from astropy import units as u

%# Declaring embedding object with specified mass of the body 
%surface_obj = SchwarzschildEmbedding(5.927e23 * u.kg)
%surface_obj.plot_hypersurface(plot_type='surface')

%# plotting the embedding hypersurface for Schwarzschild spacetime
%surface_obj.show()
%\end{minted}
\includegraphics[scale=0.5]{hypersurface_surface.png}

\section{Displaying mathematics} \label{sec:displaymath}

The most common mathematical symbols and formulas are in the amsmath
package.  \aastex\ requires this package so there is no need to
specifically call for it in the document preamble.  Most modern \latex\
distributions already contain this package.  If you do not have this
package or the other required packages, revtex4-1, latexsym, graphicx,
amssymb, longtable, and epsf, they can be obtained from 
\url{http://www.ctan.org}

Mathematics can be displayed either within the text, e.g. $E = mc^2$, or
separate from in an equation.  In order to be properly rendered, all inline
math text has to be declared by surrounding the math by dollar signs (\$).

A complex equation example with inline math as part of the explanation
follows.

\begin{equation}
\bar v(p_2,\sigma_2)P_{-\tau}\hat a_1\hat a_2\cdots
\hat a_nu(p_1,\sigma_1) ,
\end{equation}
where $p$ and $\sigma$ label the initial $e^{\pm}$ four-momenta
and helicities $(\sigma = \pm 1)$, $\hat a_i=a^\mu_i\gamma_\nu$
and $P_\tau=\frac{1}{2}(1+\tau\gamma_5)$ is a chirality projection
operator $(\tau = \pm1)$.  This produces a single line formula.  \latex\ will
auto-number this and any subsequent equations.  If no number is desired then
the {\tt\string equation} call should be replaced with {\tt\string displaymath}.

\latex\ can also handle a a multi-line equation.  Use {\tt\string eqnarray}
for more than one line and end each line with a
\textbackslash\textbackslash.  Each line will be numbered unless the
\textbackslash\textbackslash\ is preceded by a {\tt\string\nonumber}
command.  Alignment points can be added with ampersands (\&).  There should be
two ampersands per line. In the examples they are centered on the equal
symbol.

\begin{eqnarray}
\gamma^\mu  & = &
 \left(
\begin{array}{cc}
0 & \sigma^\mu_+ \\
\sigma^\mu_- & 0
\end{array}     \right) ,
 \gamma^5= \left(
\begin{array}{cc}
-1 &   0\\
0 &   1
\end{array}     \right)  , \\
\sigma^\mu_{\pm}  & = &   ({\bf 1} ,\pm \sigma) , 
\end{eqnarray}

\begin{eqnarray}
\hat a & = & \left(
\begin{array}{cc}
0 & (\hat a)_+\\
(\hat a)_- & 0
\end{array}\right), \nonumber \\
(\hat a)_\pm & = & a_\mu\sigma^\mu_\pm 
\end{eqnarray}

%% Putting eqnarrays or equations inside the mathletters environment groups
%% the enclosed equations by letter. For instance, the eqnarray below, instead
%% of being numbered, say, (4) and (5), would be numbered (4a) and (4b).
%% LaTeX the paper and look at the output to see the results.

\section{Revision tracking and color highlighting} \label{sec:highlight}

Authors sometimes use color to highlight changes to their manuscript in
response to editor and referee comments.  In \aastex\ new commands
have been introduced to make this easier and formalize the process.

The first method is through a new set of editing mark up commands that
specifically identify what has been changed.  These commands are
{\tt\string\added\{<text>\}}, {\tt\string\deleted\{<text>\}}, and
{\tt\string\replaced\{<old text>\}\{<replaced text>\}}. To activate these
commands the {\tt\string trackchanges} option must be used in the
{\tt\string\documentclass} call.  When compiled this will produce the
marked text in red.  The {\tt\string\explain\{<text>\}} can be used to add
text to provide information to the reader describing the change.  Its
output is purple italic font. To see how {\tt\string\added\{<important
added info>\}}, {\tt\string\deleted\{<this can be deleted text>\}},
{\tt\string\replaced\{<old data>\}\{<replaced data>\}}, and \break
{\tt\string\explain\{<text explaining the change>\}} commands will produce
\added{important added information}\deleted{, deleted text, and }
\replaced{old data}{and replaced data,} toggle between versions compiled with
and without the {\tt\string trackchanges} option.\explain{text explaining
the change}

>>>>

\newcommand{\sceeiitmandi}{School of Computing and Electrical Engineering, Indian Institute of Technology Mandi, India}

% \author[0000-0000-0000-0000]{The EinsteinPy Community}
% \noaffiliation

% \author[orcid-id]{name}
% \affiliation{}
% below add contributing authors (those folks that have actually worked writing the paper).
% board members get some symbol next to their names.

\author[0000-0002-0870-4665]{Shreyas Bapat}
\altaffiliation{EinsteinPy Lead Developer, EinsteinPy Board Member}
\affiliation{\sceeiitmandi}

\author[0000-0002-1900-3338]{Shilpi Jain}
\altaffiliation{EinsteinPy Board Member}
\affiliation{Department of Earth Sciences, Indian Institute of Technology Roorkee, India}

\author{Ritwik Saha}
\altaffiliation{EinsteinPy Lead Developer, EinsteinPy Board Member}
\affiliation{\sceeiitmandi}

\author{Bhavya Bhatt}
\altaffiliation{EinsteinPy Deputy Lead Developer}
\affiliation{\sceeiitmandi}

\author[0000-0003-3354-1893]{Akshita Jain}
\affiliation{\sceeiitmandi}

\collaboration{1000}{(Primary Paper Contributors)}

% Below add contributors to the EinsteinPy project

\author[0000-0003-1021-3336]{Priyanshu Khandelwal}
\altaffiliation{EinsteinPy Board Member}
\affiliation{\sceeiitmandi}

\author[0000-0002-2268-9772]{Sofía Ortín Vela}
\affil{University of Zaragoza, Spain}

@@

\author[0000-0001-7234-6148]{Varun Singh}
\affiliation{\sceeiitmandi}

@@

\author[0000-0002-5198-9010]{Manvi Gupta}
\affiliation{\sceeiitmandi}

\author{Tushar Tyagi}
\affiliation{\sceeiitmandi}

\author[0000-0003-1838-8368]{Tanmay Rustagi}
\affiliation{\sceeiitmandi}

\author[0000-0002-0757-2883]{Abhijeet Manhas}
\affiliation{\sceeiitmandi}

@@

\author[0000-0001-8302-1584]{Sashank Mishra}
\affiliation{Department of Information Technology, Indian Institute of Information Technology Allahabad}

\collaboration{1000}{(EinsteinPy Contributors)}

\correspondingauthor{Shreyas Bapat}
\email{shreyas@einsteinpy.org, bapat.shreyas@gmail.com}
